---
title: "Team Project 1"
author:
    - Pranav Manjunath (Checker, Coordinator)
    - Aiman Haider     (Presenter)
    - Xinyi Pan        (Programmer)
    - Maobin Guo       (Writer)
output:
  pdf_document: default
  word_document: default
  html_document:
    
    
    df_print: paged
---

```{r, echo=FALSE, include=FALSE, message=FALSE}

###### Clear environment and load libraries
rm(list = ls())

options(warn=-1)
options(xtable.comment = FALSE)

library(broom) 
library(ggplot2)
library(xtable)
library(rms)
library(pROC)
library(e1071)
library(caret)
require(gridExtra)
library(MASS)
library(arm)
library(dplyr)
library(tidyr)

###### Load the data
lalondedata <-
  read.table(
    "lalondedata.txt",
    header = TRUE,
    sep = ",",
    colClasses = c(
      "factor",
      "factor",
      "numeric",
      "numeric",
      "factor",
      "factor",
      "factor",
      "factor",
      "numeric",
      "numeric",
      "numeric"
    )
  )


lalondedata$diff <- lalondedata$re78 - lalondedata$re74
dim(lalondedata)
str(lalondedata)
summary(lalondedata)

lalondedata$agec <- c(scale(lalondedata$age, scale = F))
lalondedata$educc <- c(scale(lalondedata$educ, scale = F))
lalondedata$educc2 <- lalondedata$educc ^ 2

```

## Summary

Through this report we try to analyze the impact of a training program by gauging its effect on the real annual earnings and through its likeliness to help the participants earn a non-zero wage. To do so, we use  linear  and logitic regression models  respectively. We use a dataset from the National Supported Work (NSW) Demonstration (1975-1978) for the same and build models using the AIC/BIC criterion based on the statistical significance and reasonableness. We then find out that ......have significant association with annual incomes suggesting that taking the training might be an important factor associated with increase in wages. Also, We find that non-zero wages are associated with ...... suggesting that training is not a very significant factor associated with non-zero wages. Thus, it can be understood that training program as a factor does not seem to have a very strong association with improvement in wages or by providing a source of earning.


For detailed information on this research, please check the following papers. 

- [Paper 1](https://www.jstor.org/tc/accept?origin=%2Fstable%2Fpdf%2F1806062.pdf)
- [Paper 2](https://uh.edu/~adkugler/Dehejia&Wahba_JASA.pdf)


## Introduction

In the 1970s, researchers in the United States ran several randomized experiments to evaluate public policy programs. One of the most famous experiments is the National Supported Work (NSW) Demonstration, in which researchers wanted to assess whether or not job training for disadvantaged workers had an effect on their wages. Based on a subset of the investigation, in order to undersatnd the impact of the training program we need to look at two main questions:

Part I: Is there evidence that workers who receive job training tend to earn higher wages than workers who do not receive job training?

To address this question we need to look at quantify the effect of the treatment, that is, receiving job training, on real annual earnings and understanding the likely range for the effect of training. We also need to check if there is any evidence that the effects differ by demographic groups and if there are other interesting associations with wages.

Part II: Is there evidence that workers who receive job training tend to be more likely to have positive (non-zero) wages than workers who do not receive job training?

To understand this question we need to quantify the effect of the treatment, that is, receiving job training, on the odds of having non-zero wages and what would be the likely range for the effect of training. We also need to see if there is any evidence that the effects differ by demographic groups and also if there are other interesting associations with positive wages.

These questions would help us understand the potential association of the impact of the training program on the wages. To answer these questions the report uses a linear regression on the differences in wages and a logistic regression model on the odds of getting a non-zero wage after the training. It begins with an EDA of the data, tries building a model by exploring models built with the help of AIC and BIC criteria using forward and stepwise model building and chooses the most suitable model on the basis of accuracy and plausibility to answer the above questions.

# PART I

## DATA

The data we have considered for the analysis consists of 614 observations and 11 variables viz.  X (Participant ID), treat, age, educ (education level), black, hispan (Hispan), married, nodegree, re74 (real earnings in 1974), re75 (real earnings in 1975), and re78 (real earnings in 1978). Further,  the dataset contains 185 male participants that have attended the training program and 429 male participants that have not attended the training program. There are 243 black and 72 hispanic participants. 
We converted treat, black, hispan, married, nodegree into binary factor variables, educ to a discrete variable and age as a continuous variable. The columns X and re75 will not be used in this analysis. We will not use re75 as it cannot be considered as a constant baseline among the participants. 

### Exploratory Data Analysis

We initially noticed that re78 was not normally distributed so that could not be used as the response variable. The difference between re78 and re74 seemed to portray more characteristics of normal distribution hence we have taken this difference as the model’s response variable. 

```{r, echo=FALSE, out.height='20%', fig.align='center', fig.show='hold'}

ggplot(lalondedata, aes(x = diff)) +
  geom_histogram(
    aes(y = ..density..),
    color = "black",
    linetype = "dashed",
    fill = rainbow(35),
    binwidth = 2500
  ) +
  geom_density(alpha = .25, fill = "lightblue") +
  scale_fill_brewer(palette = "Blues") +
  labs(title = "Distribution of Real Annual Earnings Difference between 1978 and 1974",
       x = "Real Annual Earnings Difference") +
  theme_classic() + theme(legend.position = "none")


```

It is observed the variable age shows a slightly decreasing linear relationship with the wage differences, while variable education shows a non-linear trend. Further, we also see that categorical variables treat, black, hispan and married seem to show some differences based on the categories. We then look into the interactions and find that the association between wage differences and treat seems to differ by age; while one has an increasing trend the other has a decreasing trend. The trend between the two also shows differences by races , educ and married. We also find that trends between the wage differences and education differ by married. This is also found so for the age variable to differ by married.



```{r, echo=FALSE, out.width=c('50%', '50%'), out.height='20%', fig.show='hold', fig.align='center',message=FALSE}

# educ
ggplot(lalondedata, aes(x = educ, y = diff)) +
  geom_point(alpha = .5, colour = "blue4") +
  geom_smooth(col = "red3") +
  theme_classic() +
  labs(title = "Difference in Earnings vs Education", x = "Education",
       y = "Difference in Earnings")

ggplot(lalondedata, aes(x = age, y = diff)) +
  geom_point(alpha = .5, colour = "blue4") +
  geom_smooth(method = "lm", col = "red3") +
  theme_classic() +
  labs(title = "Difference in Earnings vs Age", x = "Age",
       y = "Difference in Earnings")

```


Unlike the relationship between age and 'treat' (Right), the relationship between education(left) and 'treat' is not linear. It indicates that there is some non-linear transformation should be performed on education. After evaluation, we decide to use the square to transform the variable.


```{r, echo=FALSE,  fig.show='hold', out.height='20%', fig.align='center' , message=FALSE}

ggplot(lalondedata, aes(x = age, y = diff)) +
  geom_point(alpha = .5, colour = "blue4") +
  geom_smooth(method = "lm", col = "red3") +
  theme_classic() +
  labs(title = "The difference in Earnings vs Age influenced by treat", x = "Treat",
       y = "Difference in Earnings") +
  facet_wrap( ~ treat)

```

The plot above illustrates the interaction between age and treat with the response variable. The annual earnings trend with the increase of age is different according to treat. Further investigation of this interaction would be exerted in the model fitting step.

## Model


### Model selection

Based on these, we try building a model using the AIC forward/ BIC and selecting the model. 

Null Model contained only the treat variable while the full model contained all the predictors along with its interactions. The null model has an AIC value of 12771.89 while the full model has an AIC value of 12764.68. 

The output of Forward AIC model is:

diff ~ treat + agec + married + educc2 + treat:agec + 
    agec:married

Where agec - age centered variable and educc - education centered variable.

The AIC model contains the variables, treat, agec, married, educc2 (squared term of educ) along with the interaction between treat & agec and agec & married.

On comparing the model to the Null Model and Full Model, we find that the above model is more statistically significant (p-value: 6.668e-07 (null) and p-value:0.3817 (full)). Further, we test excluding and including some variables and found this to indeed be the most statistically significant one. However, we add the variables black and hispan as is required for analysing the question of interest. We also add the variable educc as we retain its second order term [^Shows a significant p-value for the dataset.]

Interaction between "married" and "age" was found in this step. It's p-Value is slightly above 0.05, but it small than 0.1. The ANOVA test also indicates that it would improve our model significantly. Hence it was preserved in our final model. 


Intuitively, education duration has a strong correlation with a high school degree.  In this dataset, the correlation of the two variables is `r round(cor(as.numeric(lalondedata$nodegree), lalondedata$educ), 2)` which suggests that we could not include both of them in our model. After evaluation, education duration was preserved since it can provide more information than the high school degree variable.



### Final model

$$
\begin{aligned}
diff_{i} = &\beta_{0} + \beta_{1}*black_{i} + \beta_{2}*hispan_{i} + \beta_{3}*agec_{i} + \\ 
                 &\beta_{4}*married_{i} + \beta_{5}*treat_{i}:agec_{i} + \beta_{6}*educc_{i} + \\
                 &\beta_{7}*educc_{i}^2 + \beta_{8}agec:married_{i} + \varepsilon_{i}
\end{aligned}
$$


### Model Summary & CI

```{r echo = FALSE, results='asis' ,  fig.show='hold'}

final_model <-
  lm(
    diff ~ treat + black + hispan + agec + married + treat:agec + educc + educc2
    + agec:married,
    data = lalondedata
  )

fsta <- summary(final_model)$fstatistic
model_pvalue <- pf(fsta[1],fsta[2],fsta[3],lower.tail=F)
```

```{r echo = FALSE, results='asis',  fig.show='hold'}
xtable(final_model, 
      caption = "Coefficient-Level Estimates")

tdf <- data.frame(
    pValue = formatC(model_pvalue, format = "e", digits = 2),
    RSquare = c(round(summary(final_model)$r.squared, 2))
)

xtable(tdf, caption = "Evaluationl")

#xtable(confint(final_model), caption = "Confidence Interval")

```

The baseline values incorporated in the intercept are treat=0,black=0,hispan = 0,age = 27, and married=0. 
From the above model summary, we find that treat, agec, married, educc2 (square term of education) and the interactions of age and treat are strongly associated (statistically significant) with the differences in wages at the 0.05 level. The interaction between age and married is significant at the 0.1 significance level. With demographic factor, we notice that race (black or hispanic) is not significant with the difference in wages, p value greater than 0.05. Keeping other variables constant,
1. If the participant takes part in the training program, their annual earning is likely to increase by $813.5
2. A unit increase of age from the mean value (27 years) tend to decrease the annual earnings by $55.02
3. If the participant is married, his annual earnings tend to decrease by $469.90

(INTERPRET EDUC^2 and INTERACTION terms)


Further at a level of significance, it can be observed that …. Increase or decrease . with ….. Keeping other variables constant.

From the above, we find that treat is significant at level …… And the likely range of wage difference is ….. This suggests that training does have an association with higher wages. 

With demographic factors…..


\hfill\break

### Model Verification 

#### Residuals

```{r echo = FALSE, out.width=c('50%', '50%'), out.height='20%', fig.align='center', fig.show='hold'}

plot(final_model, which=1)
plot(final_model, which=3)

```

- The residuals are scattered randomly; there is no apparent trend in the plots.
- The error is no correlation of error terms in the plot.
- The variance of the error is constant, there is no apparent change along the x-axis. 

**Summary:** 
According to residual analysis, there is no obvious evidence indicate the assumptions of linear regression were broken. 

#### Outliers and High Leverage

\hfill\break

```{r echo = FALSE, out.width=c('30%', '30%', '30%'),  out.height='20%', message=FALSE, hold_position=TRUE}

# Outliers
plot(final_model, which=2)

# Outliers
lev_scores <- hatvalues(final_model)
p<- 9
n <- nrow(lalondedata)
plot(lev_scores, col=ifelse(lev_scores > (2*p/n), 'red2', 'navy'), type="h", ylab="Leverage score",xlab="Index",main="Leverage Scores for all observations")

# High Influence
plot(final_model, which=5, main = "Cook's Distance Analysis")

```

- There are a few outliers under this model.
- There are some high leverage points. 
- According to cook's distance, there is no high influence points (> 0.5).

**Summary:** 
There are some outliers and high leverage points; however, there are not high influence data. Hence, these data points can be preserved in the model without worry. 

#### Collinearity 

\hfill\break
```{r echo = FALSE, results='asis', message=FALSE}

xtable(tidy(vif(final_model)), caption = "VIF")

```

- According to VIF table, there is obvious colineary problem in this model. 

## Conclusion

1. Tread has positive effects on workers' annual salary because its p-value is significant. Controlling other factors, taking job training would increase $3254 on annual salary on average. It's 95% CI  is (1516, 4991)

2. The effect varies by age. The interaction of treat and age is significant in our model. Workers who received training would receive $124 per year for per 1-year increase in age, while the no-training workers' salary would decrease by $322 per year for per 1-year increase in age.

3. Other interesting associations with wages:

- Marriage would significantly bring down workers' annual salary by $1879 (95% CI is 452, 3307)

- Education duration would increase workers' salaries. For 1 unit increase for its square, the annual salary would increase by $55 (95% CI: 3, 108) 

- Age and married have weak interaction. Controlled other factors, for the married workers, while the old ones would receive more salary. One year increase on age would raise the workers' salary by $137 per year (95% CI: -2, 278)


## Deficiency

1. The final model's R-squared is only 0.088, which is relatively low. 

2. Some outliers deserve further investigation. 


\newpage

# Part II

## DATA

To answer the questions of interest, we converted the re78 variable into a binary response variable. A 0 value of the variable was taken to indicate that the participant made no earnings in 1978 while a 1 to indicate some (non-zero) monetary earnings in 1978. In other words, the variable is an indicator of the employment status of the participants in 1978 and is made our response variable. The variable re74 (real earnings in 1974) has also been removed from the analysis as the main aim was to look at the impact of training on positive wages (employment status). The other variables have been retained as was also mentioned in Part 1. 

### EDA



```{r, echo=FALSE, message=FALSE}

rm(list = ls())

lalondedata <-
  read.table(
    "lalondedata.txt",
    header = TRUE,
    sep = ",",
    colClasses = c(
      "factor",
      "factor",
      "numeric",
      "numeric",
      "factor",
      "factor",
      "factor",
      "factor",
      "numeric",
      "numeric",
      "numeric"
    )
  )

lalondedata$earn <- ifelse(lalondedata$re78 > 0, 1, 0)

lalondedata$earnf <-
  factor(
    ifelse(lalondedata$re78 > 0, 1, 0),
    levels = c(0, 1),
    labels = c("Zero", "Positive")
  )


lalondedata$agec <- lalondedata$age - mean(lalondedata$age)
lalondedata$educc <- lalondedata$educ - mean(lalondedata$educ)
lalondedata$agec2 <- lalondedata$agec ^ 2

```


We then look at some of the trends between the other variables and the response variable. On some analysis, we find that this employment variable seems to show relation with age.


```{r echo = FALSE, results='asis', out.width=c('50%', '50%'),  out.height='20%'}

ggplot(lalondedata, aes(x = earnf, y = age, fill = earnf)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette = "Reds") +
  labs(title = "Had salaries or not vs Age by Treat",
       x = "Had salaries or no?", y = "Age") +
  theme_classic() + theme(legend.position = "none") +
  facet_wrap(~ treat)

ggplot(lalondedata, aes(x = earnf, y = age, fill = earnf)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette = "Reds") +
  labs(title = "Had salaries or not vs Age by High school degree",
       x = "Had salaries or no?", y = "Age") +
  theme_classic() + theme(legend.position = "none") +
  facet_wrap(~ nodegree)

```


As reported by the first plot (left), the relation between age and employment in 1978 is different based on whether or not the employee took the training program.
The second plot points out that the relation between employment and age is also influenced by whether or not the worker has a high school degree




```{r echo = FALSE, results='asis', message=FALSE}

hispan0<-lalondedata %>%
  filter(hispan==0)
hispan1<-lalondedata %>%
  filter(hispan==1)

non_hispan_marriage <- apply(table(hispan0[,c("earnf","married")])/sum(table(hispan0[,c("earnf","married")])), 2,function(x) round(x/sum(x), 2))
hispan_marriage <- apply(table(hispan1[,c("earnf","married")])/sum(table(hispan1[,c("earnf","married")])), 2,function(x) round(x/sum(x),2))

xtable(non_hispan_marriage, caption = "Employed rate with marital status - Non Hispanic")
xtable(hispan_marriage, caption = "Employed rate with marital status - Hispanic")

```

We also observe through the two tables shown that there are some differences in the relationship between marital status and positive wages according to race. For the Hispanic participants, the unmarried workers’ unemployment rate seems to be lower than single workers of other races (8% vs. 26%). At the same time, this advantage disappears in the married Hispanic workers. The married Hispanic workers’ unemployment rate is 28%, while Non-Hispanic races’ married unemployment rate is 21%. This interesting sign of interaction would be further investigated in the model fitting.^[ The Chi-square tests for independence are not significant. The variables are explored later.]

## Model

### Model selection

Based on the above trends, we build a logistic regression model using the AIC forward model selection method. For doing so we construct our 
Null Model as:

$$
\begin{aligned}
logit(\pi_{i}) = &\beta_{0} + \beta_{1}*treat_{i}
\end{aligned}
$$

Full Model as:

$$
\begin{aligned}
logit(\pi_{i}) = &\beta_{0} + (\beta_{1}*treat_{i} + \beta_{2}*black_{i} + \beta_{3}*hispan_{i} + \beta_{4}*agec_{i} + \beta_{5}*married_{i} + \beta_{6}*nodegree_{i})^2 + \beta_{5}*agec_{i}^2 + \varepsilon_{i}
\end{aligned}
$$
[^ A number of extra variables are considered to understand their role in the model and get to the most significant variables among these.]


1. On the basis of the above, the logistic regression model is developed after considering the stepwise and forward built models using the AIC criteria and then selected. The Null Model used is :
$$
\begin{aligned}
logit(\pi_{i}) = &\beta_{0} + \beta_{1}*treat_{i} + \beta_{2}*black_{i} + \beta_{3}*agec_{i} + \beta_{4}*agec_{i}^2 + \varepsilon_{i}
\end{aligned}
$$

and the Full Model used is:

$$
\begin{aligned}
logit(\pi_{i}) = &\beta_{0} + (\beta_{1}agec_{i} + \beta_{2}educc_{i} + \beta_{3}treat_{i} + \beta_{4}black_{i} + \beta_{5}hispan_{i} + \beta_{6}married_{i} + \beta_{7}nodegree_{i})^2+ \\ &\beta_{8}*age_{i}c^2 + \varepsilon_{i}
\end{aligned}
$$



Using the AIC forward model, we find that the square transform of “age” is significant in the model. After verification with ANOVA test, we decide to keep this transformation in the final model

As shown in the model output, the interactions between “age” &“nodegree” and “hispan” & “married” seems to be statistically significant, with a p value of 0.03 and 0.01 respectively. Interestingly, even though the interaction between treat and age is weakly significant (p-value=0.09), the ANOVA test confirmed its significance and hence we have included this in the final model. As the variable treat is statistically not significant (p-value=0.4), we preserve it in our final model as it is used to answer the questions of interest.




### Final model


$$
\begin{aligned}
logit(\pi_{i}) = &\beta_{0} + \beta_{1}*treat_{i} + \beta_{2}*black_{i} + \beta_{3}*agec_{i} + \beta_{4}*agec_{i}^2 + \beta_{5}*nodegree_{i} + \beta_{6}*hispan_{i} +\\                          &\beta_{7}*married_{i} + \beta_{8}*agec_{i}:nodegree_{i} + \beta_{9}*hispan_{i}:married_{i} + \beta_{10}*agec_{i}:treat_{i}+ \varepsilon_{i}
\end{aligned}
$$


```{r, echo=FALSE, message=FALSE}

full_model <-
  glm(
    earn ~ treat + black + agec + agec2  
           + nodegree + hispan + married 
           + nodegree:agec + hispan:married + agec:treat,
    
    data = lalondedata,
    family = binomial
  )

rawresid <- residuals(full_model,"resp")

model_sum <- summary(full_model)

```

### Model Summary & CI


```{r echo = FALSE, results='asis', message=FALSE}

xtable(model_sum, caption = "Coefficients")
 
model_confint <- confint(full_model)

xtable(model_confint, caption = "Confidence Interval")

tdf <- data.frame(
    NULL_deviance = round(model_sum$null.deviance, 2),
    Residual_deviance = round(model_sum$deviance, 2)
)

xtable(tdf, caption = "Deviance")

```

The baseline values taken in the intercept is treat=0 and black=0, nodegree=0, hispan=0, age=27, and married=0. Keeping all predictors at 0, the birth weight of the newborn would be 46.39 ounces (does not make any sense as centering is not done).
Keeping other variables constant,
1. The odds ratio of a black employee having earnings in 1978 decreases by 0.417 times when compared to a non-black employee.
2. The odds ratio of a hispanic employee having earnings in 1978 increases by 3.32 times when compared to a non-hispanic employee.






### Model Verification

#### Residuals

```{r echo = FALSE, results='asis', out.width=c('30%', '30%', '30%'), fig.align='center', out.height='25%'}

binnedplot(
  x = fitted(full_model),
  y = rawresid,
  xlab = "Pred. probabilities",
  col.int = "red4",
  ylab = "Avg. residuals",
  main = "Binned residual plot",
  col.pts = "navy"
)

#####
binnedplot(
  x = lalondedata$agec,
  y = rawresid,
  xlab = "Age (Centered)",
  col.int = "red4",
  ylab = "Avg. residuals",
  main = "Binned residual plot",
  col.pts = "navy"
)

binnedplot(
  x = lalondedata$educc,
  y = rawresid,
  xlab = "Education (Centered)",
  col.int = "red4",
  ylab = "Avg. residuals",
  main = "Binned residual plot",
  col.pts = "navy"
)

```

The binned plot above illustrates the average residuals vs predicted probabilities. As shown, more than 95% of points reside inside the red band (within 95% confidence intervals) and no distinct pattern is observed. The binned plots for average residuals vs individual variables have been plotted and these plots also do not violate the above rules. This concludes a strong justification for the model’s efficiency.

Null model deviance : 666.5
Final model deviance : 627.54
The decrease of deviance indicates that the model is valid.


#### Collinearity


```{r echo = FALSE, fig.show='hold', results='asis', message=FALSE}

xtable(tidy(vif(full_model)), caption = "VIF")

```

According to the table, most of the VIF values for the predictors are below 5, indicating moderately correlated. However, age centered seems to have a VIF value of 6.8, indicating high correlation. 

#### Summary 

Various indicators point out that the final model is valid and can be used to answer the questions  about the dataset. 

### Model Assessment

```{r echo = FALSE, results='asis', message=FALSE}

Conf_mat <-
  confusionMatrix(as.factor(ifelse(
    fitted(full_model) >= mean(lalondedata$earn), "1", "0"
  )),
  as.factor(lalondedata$earn), positive = "1")

                
xtable(Conf_mat$table, caption = "Confusion Matrix")

```


The Sensitivity value, also known as true positive rate (TP/TP+FN){Latex} of the model is 0.59.

The Specificity value, also known as true negative rate (TN/FP+TN){Latex} of the model is  0.64

The Accuracy (TP+TN/TP+TN+FP+FN) {Latex}of the model is 0.60, 60%.


#### ROC

```{r echo = FALSE, results='asis', fig.align='center', message=FALSE,  out.height='25%'}


cutoff = "best"

roc <- roc(lalondedata$earn,
           fitted(full_model),
           plot=T,
           print.thres=cutoff,
           legacy.axes=T, 
           col="blue2")

roc_value <-coords(roc, cutoff, transpose = FALSE)

```

The ROC curve above has a AUC value of 0.65 and a cut-off value of 0.81.

Discussion

1. Is there evidence that workers who receive job training tend to be more likely to have positive (non-zero) wages than workers who do not receive job training? Quantify the effect of the treatment, that is, receiving job training, on the odds of having non-zero wages.LIMITS

According to the model, the variable treat is not statistically significant as it has a p value greater than 0.05. Hence receiving job training would not directly influence the odds of these worker’s positive wages and cannot quantify the effect of treatment on the odds of having non-zero wages. However, the interaction item “treat1:agec” is relatively significant. {talk later}


3) Is there any evidence that the effects differ by demographic groups?

Yes. In the logistic regression model, black and hispan are statistically significant. For a black worker, his odds of getting a job are 41% lower than a non-black worker under the same other conditions.
The odds ratio of a hispanic employee having positive earnings in 1978 increases by 3.32 times when compared to a non-hispanic employee.


4) Are there other interesting associations with positive wages that are worth mentioning?

Race of Hispanic and marital status. Controlling other factors, a Hispanic workers odd to be employed is 231% higher than others. However, the effect would bring down by being married. A married Hispanic worker’s odds of getting a job are 45% lower than others under the same conditions.

Age. Age is a complex factor in our model. Generally speaking, aging would increase these workers’ odds of being jobless. Taking the job training may slow down the trends.



## Conclusion


To understand the impact of job training on positive wages, we built a logistic regression model. We confirmed the model using AIC forward selection and ANOVA tests. The assessment of the model was done using binned residual plots and identifying specificity, sensitivity, accuracy, and area under the ROC. We noticed that the variable treat is not statistically significant, inferring that job training would not directly influence the odds of these worker’s positive wages. Race and Age seemed to be statistically significant when predicting the positive wages. Three interactions, Age & treat, hispan & married, age & "no degree" are also statistically significant as determined by p value less than 0.05. The effect of age is hard to interpret because it involves square transformation and interaction with treat.


## Limitation 

Sensitivity and accuracy are relatively low.
Data on the long term effect of training is missing in this dataset. Intuitively, job training would exert its influence on people’s work in the long run. However, it can not be verified in this analysis. There is an imbalance in 
The number of participants who joined the training program
The ratio of hispanics to non-hispanics



\newpage
# Appendix I (Part1 R Code)


````
`r paste(readLines('Part 1.R'), collapse = '\n')`
````


\newpage
# Appendix II (Part2 R Code)

````
`r paste(readLines('Part 2.R'), collapse = '\n')`
````





















































































